```{r}
library(censusapi)
library(tidyverse)
```


INSPECT METADATA:
```{r}
apis <- listCensusApis()
#View(apis)
```

```{r}
acs_vars <- listCensusMetadata(
    name = "acs/acs1", 
    type = "variables",
    vintage=2019)
#View(acs_vars)
```

```{r}
acs_geogs <- listCensusMetadata(
    name = "acs/acs1", 
    type = "geographies",
    vintage=2019)
#View(acs_geogs)
```

```{r}
acs_group <- listCensusMetadata(
    name = "acs/acs1", 
    type = "group",
    vintage=2019)
#View(acs_group)
```



BEGIN IMPORT:
```{r}
#Import Race/Ethnicity for Nashville MSA
acs_race <- getCensus(
  key = CENSUS_KEY,  
  name = "acs/acs1", 
  vintage = 2019, 
  vars = c("group(B03002)"), 
  region = "metropolitan statistical area/micropolitan statistical area:34980")
```

```{r}
#Import Educational Attainment for Nashville MSA
acs_edu <- getCensus(
  key = CENSUS_KEY,  
  name = "acs/acs1", 
  vintage = 2019, 
  vars = c("group(B15002)"), 
  region = "metropolitan statistical area/micropolitan statistical area:34980")
```

```{r}
#Import Age for Nashville MSA
acs_age <- getCensus(
  key = CENSUS_KEY,  
  name = "acs/acs1", 
  vintage = 2019, 
  vars = c("group(B01001)"), 
  region = "metropolitan statistical area/micropolitan statistical area:34980")
```

IMPORT LABELS:
```{r}
race_labels <- acs_vars %>%
  filter(group == 'B03002') %>%
  select(c('name', 'label'))
```

```{r}
edu_labels <- acs_vars %>%
  filter(group == 'B15002') %>%
  select(c('name', 'label'))
```

```{r}
age_labels <- acs_vars %>%
  filter(group == 'B01001') %>%
  select(c('name', 'label'))
```

REMOVE UNNEEDED COLUMNS:
```{r}
acs_race <- acs_race %>%
  select(c(ends_with('e')))
acs_edu <- acs_edu %>%
  select(c(ends_with('e')))
acs_age <- acs_age %>%
  select(c(ends_with('e')))
```

CLEAN LABELS:
```{r}
race_labels <- race_labels %>% mutate(label = str_replace_all(string = label, pattern = 'Estimate!!Total:!!', replacement = ''))
edu_labels <- edu_labels %>% mutate(label = str_replace_all(string = label, pattern = 'Estimate!!Total:!!', replacement = ''))
age_labels <- age_labels %>% mutate(label = str_replace_all(string = label, pattern = 'Estimate!!Total:!!', replacement = ''))
race_labels <- race_labels %>% mutate(label = str_replace_all(string = label, pattern = '!!', replacement = ''))
edu_labels <- edu_labels %>% mutate(label = str_replace_all(string = label, pattern = '!!', replacement = ''))
age_labels <- age_labels %>% mutate(label = str_replace_all(string = label, pattern = '!!', replacement = ''))
```

REARRANGE LABELS TO MATCH CENSUS IMPORT ORDER:
```{r}
race_labels <- arrange(race_labels, name)
edu_labels <- arrange(edu_labels, name)
age_labels <- arrange(age_labels, name)
```

CREATING LOOKUP VECTOR:
```{r}
lookup_vector_race <- race_labels$label
lookup_vector_race <- append(lookup_vector_race, "Location")
lookup_vector_edu <- edu_labels$label
lookup_vector_edu <- append(lookup_vector_edu, "Location")
lookup_vector_age <- age_labels$label
lookup_vector_age <- append(lookup_vector_age, "Location")
```

REPLACING LABELS:
```{r}
rownames(race_labels) <- NULL
rownames(edu_labels) <- NULL
rownames(age_labels) <- NULL
```

```{r}
colnames(acs_race) <- lookup_vector_race
colnames(acs_edu) <- lookup_vector_edu
colnames(acs_age) <- lookup_vector_age
```

VIEW DATAFRAMES
```{r}
#View(acs_age)
#View(acs_edu)
#View(acs_race)
```

MATH OUT THE COLUMNS TO GET CORRECT CATEGORIES, THEN CLEAN TO GET FINAL DATAFRAMES
START WITH AGE DATA
```{r}
#colnames(acs_age)
```

SUBSET TO MALE AND FEMALE, PREPARE TO CONCATENATE
```{r}
#Pull out columns with Female Data
age_female <- acs_age %>%
  select(c(contains('Female')))
#Rename to column names that will match names for Male Data
names(age_female) = sub('Female', 'Total', names(age_female))
#Add row column to mark gender, for reference. Will delete later.
age_female$gender = 'Female' 

#Follow same process, but for Male Data
age_male <- acs_age %>%
  select(-c('EstimateTotal:', 'Location', contains('Female')))
age_male$gender = 'Male'
names(age_male) = sub('Male', 'Total', names(age_male))
```

CONCATENATE MALE AND FEMALE
```{r}
age_df <- rbind(age_male, age_female)
```

SUM ROWS TO CORRECT CATEGORIES, REMOVE UNNEEDED ROWS:
```{r}
age_rows <- c('Male', 'Female')
#Add columns to match sample corporation data, then remove original columns
age_df <- age_df %>%
  mutate(
    age_under_20_years = rowSums(age_df[, c(2:6)]),
    age_20_29_years = rowSums(age_df[, c(7:10)]), 
    age_30_39_years = rowSums(age_df[, c(11:12)]),
    age_40_49_years = rowSums(age_df[, c(13:14)]),
    age_50_59_years = rowSums(age_df[, c(15:16)]),
    age_60_years_and_over = rowSums(age_df[, c(17:24)]),
    ) %>%
  select(-c(2:24))
row.names(age_df) <- age_rows
```

```{r}
age_df <- age_df %>%
  select(-gender) 
```

CREATE A ROW FOR TOTALS, CONCATENATE TO AGE_DF
```{r}
Total_age <- data.frame(colSums(age_df[,]))
```

```{r}
Total_age <- data.frame(t(Total_age))
```

```{r}
colnames(Total_age) <- colnames(age_df)
age_df <- rbind(age_df, Total_age)
```

RENAME AND REORDER ROWS
```{r}
age_df$order <- c(2,3,1)
age_df <- age_df %>% 
  arrange(order) %>%
  select(-order)
rownames(age_df) <- c('Total', 'Male', 'Female')
```

```{r}
#View(age_female)
#View(age_male)
#View(age_df)
```

MATH OUT THE COLUMNS TO GET CORRECT CATEGORIES, THEN CLEAN TO GET FINAL DATAFRAMES
CONTINUE WITH EDUCATION DATA

SUBSET TO MALE AND FEMALE, PREPARE TO CONCATENATE
```{r}
#Follow same steps as above, with age data
edu_female <- acs_edu %>%
  select(c(contains('Female')))
names(edu_female) = sub('Female', 'Total', names(edu_female))
edu_female$gender = 'Female'
edu_male <- acs_edu %>%
  select(-c('EstimateTotal:', 'Location', contains('Female')))
edu_male$gender = 'Male'
names(edu_male) = sub('Male', 'Total', names(edu_male))
```

CONCATENATE MALE AND FEMALE
```{r}
edu_df <- rbind(edu_male, edu_female)
```

SUM ROWS TO CORRECT CATEGORIES, REMOVE UNNEEDED ROWS:
```{r}
edu_rows <- c('Male', 'Female')
edu_df <- edu_df %>%
  mutate(
    no_high_school_diploma = rowSums(edu_df[, c(2:9)]),
    high_school_graduate = edu_df[10], 
    some_college_no_degree = rowSums(edu_df[, c(11:12)]),
    associates_degree = edu_df[13], 
    bachelors_degree = edu_df[14], 
    masters_degree = edu_df[15], 
    professional_degree_school = edu_df[16],
    doctorate_degree = edu_df[17]
    ) %>%
  select(-c(2:17))
row.names(edu_df) <- edu_rows
```


```{r}
edu_df <- edu_df %>%
  select(-gender) 
```

CREATE A ROW FOR TOTALS, CONCATENATE TO AGE_DF
```{r}
Total_edu <- data.frame(colSums(edu_df[,]))
```

```{r}
Total_edu <- data.frame(t(Total_edu))
```

```{r}
colnames(Total_edu) <- colnames(edu_df)
edu_df <- rbind(edu_df, Total_edu)
```

RENAME AND REORDER ROWS
```{r}
edu_df$order <- c(2,3,1)
edu_df <- edu_df %>% 
  arrange(order) %>%
  select(-order)
rownames(edu_df) <- c('Total', 'Male', 'Female')
```

```{r}
#View(edu_female)
#View(edu_male)
#View(edu_df)
#View(acs_edu)
```


CONTINUE CLEANING WITH RACE DATA SET

SUBSET TO NOT HISP/LAT VS HISP/LAT, PREPARE TO CONCATENATE
```{r}
#Follow same steps as above, but now splitting data between Hispanic/Latino and Non
race_not <- acs_race %>%
  select(c(contains('Not')))
names(race_not) = sub('Not Hispanic or Latino:', '', names(race_not))
race_not$race = 'Not Hispanic or Latino' 
race_hisp <- acs_race %>%
  select(-c('EstimateTotal:', 'Location', contains('Not')))
race_hisp$race = 'Hispanic or Latino'
names(race_hisp) = sub('Hispanic or Latino:', '', names(race_hisp))
```

```{r}
#renaming first column to total
colnames(race_not)[colnames(race_not) == ''] <- 'Total'
colnames(race_hisp)[colnames(race_hisp) == ''] <- 'Total'
```

CONCATENATE MALE AND FEMALE
```{r}
race_df <- rbind(race_not, race_hisp)
```

```{r}
#View(race_df)
```

Renaming Index:
```{r}
race_rows <- c('Not Hispanic or Latino', 'Hispanic or Latino')
row.names(race_df) <- race_rows
```

GETTING RID OF RACE COLUMN:
```{r}
race_df <- race_df %>%
  select(-race) 
```

CREATE A ROW FOR TOTALS, CONCATENATE TO AGE_DF
```{r}
Total_race <- data.frame(colSums(race_df[,]))
```

```{r}
Total_race <- data.frame(t(Total_race))
```

```{r}
colnames(Total_race) <- colnames(race_df)
race_df <- rbind(race_df, Total_race)
```

RENAME AND REORDER ROWS
```{r}
race_df$order <- c(2,3,1)
race_df <- race_df %>% 
  arrange(order) %>%
  select(-order)
rownames(race_df) <- c('Total', 'Not Hispanic or Latino', 'Hispanic or Latino')
```

```{r}
census_df <- list(age_df, race_df, edu_df)
```



